---
title: "Samson_HW5"
author: "Danielle Samson"
date: "2025-12-01"
output:
  html_document: default
  pdf_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning= FALSE, message= FALSE, error= FALSE)
```

## Choose a city to work with
In the following chunk, I read in the data, looked through the list of potential cities then decided to work with Denver. I then filtered the dataset to include only the data from Denver and used the provided coordinates to turn the dataframe into a simple feature object. Finally, I generated the plot to help visualize where I am and where I need to go with my code
```{r choose a city}
library(tidyverse)
library(sf)
library(tigris)
library(ggthemes)
#read in the data

homicides <- read.csv("homicide-data.csv")

#choose a city

print(unique(homicides$city))

denver_homicides <- homicides %>%
  filter(city == 'Denver') %>% 
  st_as_sf(coords = c("lon", "lat"),
               crs = 4269)

#evaluate the current plot

denver_homicides %>%
  ggplot() +
  geom_sf()
```

## Input city boundaries
In this chunk, I took the Colorado geometry from the tigris package and then filtered the data to include only Denver. Looking back, I think I could have done this in one step by immediately filtering by the NAME column. Again, I checked in on the plot
```{r results='hide'}
#extract denver boundaries from tigris
places <- tigris::places(state = "CO",
                                    cb = FALSE,
                                    year = "2024")
denver_boundaries <- places %>%
  filter(NAME == "Denver")
```

```{r evaluate the current plot}
denver_boundaries %>%
  ggplot() +
  geom_sf()

```



